### 2020-05-23 presentation.tex
presentation: add skeleton presentation, adjust build

\documentclass{f4_beamer_metropolis}

\title{Realistic real-time rendering of skin under consideration of physical basis}
\subtitle{Master seminar SS2020}
\author{Dennis Grabowski, B.Sc}
\date{25.05.2020}

\bibliography{paper.bib}

\begin{document}

% Cannot be moved into the class
\begin{frame}{Table of contents}
    \tableofcontents[hideallsubsections]
  \end{frame}

\section{Prologue}
\label{sec:intro}

\begin{frame}{Why should skin be any different from other materials?}
  \begin{itemize}
    \item Good question!
  \end{itemize}

\end{frame}

\begin{frame}[standout]
  Any questions?
\end{frame}

\section{Subsurface scattering}
\label{sec:subsurface}

\section{Application of aforementioned theoretical concepts}
\label{sec:application}

\section{Evaluation and discussion}
\label{sec:evalanddiscuss}

\section{Conclusion and Outlook}
\label{sec:outlook}

\section{Literature}

\begin{frame}[allowframebreaks]{Literature}
  \printbibliography
\end{frame}

\end{document}